\section{The canonical Kalton--Peck input and transport}

\begin{definition}[The presentation used for $Z_2$]\label{def:kp-presentation}
For a real sequence $x=(x_n)_{n\geq0}$ with $\sum_n x_n^2<\infty$, put
\[
  \|x\|_2=\left(\sum_{n=0}^{\infty}x_n^2\right)^{1/2}.
\]
For a square-summable real sequence $x$, define its Kalton--Peck centralizer coordinatewise by
\[
  (\mathrm{KP}(x))_n=2x_n\log\frac{|x_n|}{\|x\|_2},
\]
with the whole coordinate declared to be zero when $x_n=0$; this convention also covers the zero
sequence.  A pair $(u,x)$ of real sequences is \emph{admissible} when both $x$ and
$u-\mathrm{KP}(x)$ are square-summable, and for such a pair put
\[
  q(u,x)=\|u-\mathrm{KP}(x)\|_2+\|x\|_2.
\]
A \emph{real Kalton--Peck presentation} of a real normed space $Z$ consists of an injective
real-linear coordinate map
\[
  \kappa_Z:Z\longrightarrow(\mathbb R^{\mathbb N})^2
\]
whose range is exactly the admissible pairs, together with constants $c,C>0$ such that
\[
  c\,q(\kappa_Zz)\leq\|z\|\leq C\,q(\kappa_Zz)
  \qquad(z\in Z).
\]
\lean{KaltonPeck.IsSquareSummable, KaltonPeck.Support.IsSquareSummable, KaltonPeck.l2Norm, KaltonPeck.Support.l2Norm, KaltonPeck.centralizer, KaltonPeck.Support.centralizer, KaltonPeck.IsAdmissiblePair, KaltonPeck.Support.IsAdmissiblePair, KaltonPeck.kaltonPeckQuasiNorm, KaltonPeck.Support.kaltonPeckQuasiNorm, KaltonPeck.RealKaltonPeckPresentation, KaltonPeck.Support.RealKaltonPeckPresentation}
\leanok
\end{definition}

\begin{lemma}[Normalization adapter for the canonical model]
\label{lem:kp-normalization}
For a square-summable real sequence $x$, put
\[
  K_0(x)_n=x_n\log\frac{|x_n|}{\|x\|_2},
\]
with value zero when $x_n=0$.  Then
\[
  \mathrm{KP}(x)=2K_0(x).
\]
Call a pair $(a,x)$ \,\emph{$K_0$-admissible} when
\[
  a\in\ell_\infty,\qquad x\in\ell_2,
  \qquad a-K_0(x)\in\ell_2.
\]
The ambient real-linear coordinate map on pairs of real sequences
\[
  e_0(u,x)=(u/2,x)
\]
restricts to a bijection from the project-admissible pairs to the $K_0$-admissible pairs, with
inverse $(a,x)\mapsto(2a,x)$.  If
\[
  q_0(a,x)=\|a-K_0(x)\|_2+\|x\|_2,
\]
then
\[
  q_0(e_0(u,x))
   =\frac12\|u-\mathrm{KP}(x)\|_2+\|x\|_2
\]
and consequently
\[
  \frac12q(u,x)\leq q_0(e_0(u,x))\leq q(u,x).
\]
\uses{def:kp-presentation}
\lean{KaltonPeck.Support.Coordinates.k0Centralizer, KaltonPeck.Support.Coordinates.IsBoundedSequence, KaltonPeck.Support.Coordinates.IsK0AdmissiblePair, KaltonPeck.Support.Coordinates.k0QuasiNorm, KaltonPeck.Support.Coordinates.normalizationEquiv, KaltonPeck.Support.Coordinates.kpNormalization}
\leanok
\end{lemma}

\begin{proof}
Let $\mathrm e=\exp(1)$.  If $x=0$, the asserted bound
$|K_0(x)_n|\leq\|x\|_2/\mathrm e$ is immediate.  Otherwise put
$r_n=|x_n|/\|x\|_2$.  At a nonzero coordinate, $0<r_n\leq1$, and the elementary inequality
$r|\log r|\leq1/\mathrm e$ gives
\[
  |K_0(x)_n|=\|x\|_2r_n|\log r_n|\leq\frac{\|x\|_2}{\mathrm e}.
\]
The zero-coordinate convention gives the same bound at the remaining coordinates.  Thus
$K_0(x)\in\ell_\infty$.  Since every square-summable sequence is bounded, the condition
$a-K_0(x)\in\ell_2$ also forces $a\in\ell_\infty$.

The identity $\mathrm{KP}=2K_0$ is coordinatewise.  Therefore
\[
  \frac u2-K_0(x)=\frac12\bigl(u-\mathrm{KP}(x)\bigr).
\]
For an admissible pair, the right-hand side is square-summable, while the preceding bound shows that
$u/2$ is bounded.  Conversely, a $K_0$-admissible pair satisfies
$2a-\mathrm{KP}(x)=2(a-K_0(x))\in\ell_2$.  This proves preservation and reflection of the
admissibility conditions, the inverse formula, and the exact quasi-norm identity.  The two
inequalities follow by comparing $\frac12A+B$ with $A+B$ for $A,B\geq0$.
\end{proof}

\begin{theorem}[Canonical Kalton--Peck Banach model]\label{thm:kp-canonical-banach}
For the exact centralizer and quasi-norm of Definition~\ref{def:kp-presentation}, the admissible
pairs are closed under coordinatewise real-linear operations, and $q$ is equivalent to a complete
norm.  Consequently there exists a complete real normed space $Z_2^{\mathrm{can}}$ whose coordinate
map is a Kalton--Peck presentation.  In this model, the finite-coordinate pairs are dense.  Fix one
such presented space for the remainder of the blueprint.  Its second-coordinate map
$p:Z_2^{\mathrm{can}}\to\ell_2$ is a bounded surjection, while
$i:\ell_2\to Z_2^{\mathrm{can}}$, $i(x)=(x,0)$, is a bounded injection satisfying
$\ran i=\ker p$.  Both $\ell_2$ and $Z_2^{\mathrm{can}}$ are infinite-dimensional.
\uses{def:kp-presentation}
\lean{KaltonPeck.Support.Coordinates.IsFiniteCoordinatePair, KaltonPeck.Support.Coordinates.CanonicalRealKaltonPeck, KaltonPeck.Support.Coordinates.instNormedAddCommGroupCanonicalRealKaltonPeck, KaltonPeck.Support.Coordinates.instNormedSpaceCanonicalRealKaltonPeck, KaltonPeck.Support.Coordinates.instCompleteSpaceCanonicalRealKaltonPeck, KaltonPeck.Support.Coordinates.canonicalRealKaltonPeckPresentation, KaltonPeck.Support.Coordinates.CanonicalL2, KaltonPeck.Support.Coordinates.canonicalL2Quotient, KaltonPeck.Support.Coordinates.canonicalL2Quotient_apply, KaltonPeck.Support.Coordinates.canonicalL2Quotient_surjective, KaltonPeck.Support.Coordinates.canonicalL2_not_finiteDimensional, KaltonPeck.Support.Coordinates.canonicalKaltonPeckBanach, KaltonPeck.Support.Symplectic.canonicalL2Inclusion, KaltonPeck.Support.Symplectic.canonicalL2Inclusion_coordinates, KaltonPeck.Support.Symplectic.canonicalL2Inclusion_injective, KaltonPeck.Support.Symplectic.canonicalRealKaltonPeck_not_finiteDimensional, KaltonPeck.Support.Symplectic.canonicalL2Inclusion_range}
\leanok
\end{theorem}

\begin{proof}
Let $\pi$ be the second-coordinate map and let $D$ be the subspace of finite-coordinate admissible
pairs.  Finite-support sequences are dense in $\ell_2$, by convergence of the coordinate partial
sums.  Since $\ker\pi$ is linearly equivalent to $\ell_2$ through the first coordinate, with a
bounded inverse, it follows that $\ker\pi\subseteq\overline D$.

The section $s(x)=(\mathrm{KP}(x),x)$ satisfies $\pi s=\mathrm{id}$,
$\|s(x)\|\leq4\|x\|_2$, and belongs to $D$ whenever $x$ has finite support.  Given $z$ and
$\varepsilon>0$, choose such an $x$ with
$\|\pi z-x\|_2<\varepsilon/8$.  Put $d=s(x)$,
$c=s(\pi(z-d))$, and $k=z-d-c$.  Then $d\in D$, $k\in\ker\pi$, and
$\|c\|<\varepsilon/2$.  Choose $e\in D$ with $\|k-e\|<\varepsilon/2$.  Hence
$d+e\in D$ and
\[
  \|z-(d+e)\|\leq\|c\|+\|k-e\|<\varepsilon,
\]
which proves density.  The natural interleaved basis and its dense span agree with the source
model of Kalton--Peck \cite[Theorem~4.10]{KP}, as recalled in
Castillo--Cuellar--Gonz\'alez--Pino \cite[\S3]{CCGP}.
\end{proof}

\begin{theorem}[Equivalence of presented models]\label{thm:presentation-equivalence}
Let $X$ and $Y$ be complete real normed spaces with Kalton--Peck presentations.  There is a unique
real-linear bijection $e:X\to Y$ satisfying
\[
  \kappa_Y(ex)=\kappa_X(x)\qquad(x\in X),
\]
where $\kappa_X$ and $\kappa_Y$ are the two coordinate maps.  It is a bounded linear isomorphism.
If $a_X,b_X,a_Y,b_Y>0$ are comparison constants for the two presentations, then
\[
  \|ex\|\leq \frac{b_Y}{a_X}\|x\|,
  \qquad
  \|e^{-1}y\|\leq \frac{b_X}{a_Y}\|y\|.
\]
\uses{def:kp-presentation}
\lean{KaltonPeck.Support.Coordinates.AreComparisonConstants, KaltonPeck.Support.Coordinates.presentationEquiv, KaltonPeck.Support.Coordinates.presentationEquiv_spec}
\leanok
\end{theorem}

\begin{proof}
For $x\in X$, the pair $\kappa_X(x)$ is admissible, so surjectivity of the $Y$-presentation gives a
point $ex\in Y$ with $\kappa_Y(ex)=\kappa_X(x)$; injectivity of $\kappa_Y$ makes it unique.
Coordinate linearity and injectivity show that $e$ preserves zero, addition, and real scalar
multiplication.  The same construction from $Y$ to $X$ is inverse to $e$.  Finally,
\[
  \|ex\|\leq b_Yq(\kappa_Y(ex))=b_Yq(\kappa_Xx)\leq (b_Y/a_X)\|x\|,
\]
and the symmetric calculation gives the inverse bound.  Thus $e$ is a bounded linear isomorphism.
\end{proof}

\begin{theorem}[Canonical Kalton--Swanson form]\label{thm:ks-primary}
Let $Z_{2,0}$ denote the published canonical model based on $K_0$, with strong form
$\Omega_0$ from Castillo--Cuellar--Gonz\'alez--Pino, Proposition~3.1.  Write
\[
  D_0:Z_{2,0}\longrightarrow\strongdual{Z_{2,0}},
  \qquad (D_0z)(w)=\Omega_0(z,w),
\]
for its induced continuous linear isomorphism.  On finite coordinate vectors,
\[
  \Omega_0((a,x),(b,y))=\langle a,y\rangle-\langle b,x\rangle,
\]
and for arbitrary vectors this denotes its continuous extension, not a difference of two
separately asserted convergent series.

On the project-normalized canonical model define
\[
  \Omega_{\mathrm{can}}(z,w)=2\Omega_0(e_0z,e_0w).
\]
Thus, when $z$ and $w$ have project coordinates $(u,x)$ and $(v,y)$, respectively, the combined
finite-coordinate expression is
\[
  \Omega_{\mathrm{can}}(z,w)
    =\langle u,y\rangle-\langle v,x\rangle.
\]
Its induced map is
\[
  D_{\mathrm{can}}=2e_0^*D_0e_0,
\]
where $e_0^*f=f\circ e_0$.  Hence $\Omega_{\mathrm{can}}$ is a continuous strong alternating form.
\uses{thm:kp-canonical-banach,lem:kp-normalization,def:strong-form}
\lean{KaltonPeck.Support.Symplectic.standardBasisSequence, KaltonPeck.Support.Symplectic.canonicalFirstBasisVector, KaltonPeck.Support.Symplectic.canonicalSecondBasisVector, KaltonPeck.Support.Symplectic.canonicalBasisCoordinates, KaltonPeck.Support.Symplectic.canonicalKaltonSwansonForm, KaltonPeck.Support.Symplectic.canonicalKaltonSwansonForm_finite_coordinates}
\leanok
\end{theorem}

\begin{proof}
Proposition~3.1 of Castillo--Cuellar--Gonz\'alez--Pino \cite{CCGP} states that $\Omega_0$ is strong.
For the chosen coordinates,
\[
  2\Omega_0((u/2,x),(v/2,y))
    =\langle u,y\rangle-\langle v,x\rangle.
\]
Pullback by the bounded linear isomorphism $e_0$ preserves continuity and alternation, and
multiplication by the nonzero scalar $2$ preserves strongness.  Evaluation gives the displayed
formula for $D_{\mathrm{can}}$.
\end{proof}

\begin{theorem}[Transported Kalton--Swanson form]\label{thm:ks-transport}
Let $X$ be a complete real normed space with a Kalton--Peck presentation, and let
$e:X\simeq Z_2^{\mathrm{can}}$ be the coordinate-matching bounded linear isomorphism.  Then
\[
  \Omega_X(x,y)=\Omega_{\mathrm{can}}(ex,ey)
\]
is a continuous strong alternating form on $X$.  If
$e^*:\strongdual{Z_2^{\mathrm{can}}}\to\strongdual X$ denotes pullback, $e^*f=f\circ e$, then the
induced isomorphism is
\[
  D_X=e^*\circ D_{\mathrm{can}}\circ e:X\longrightarrow\strongdual X.
\]
\uses{thm:ks-primary,thm:presentation-equivalence,def:strong-form}
\lean{KaltonPeck.Support.Symplectic.transportedKaltonSwansonForm, KaltonPeck.Support.Symplectic.transportedKaltonSwansonForm_apply}
\leanok
\end{theorem}

\begin{proof}
Composition with $e$ in both variables preserves bilinearity, continuity, and alternation.  For
$x,y\in X$,
\[
  (e^*D_{\mathrm{can}}ex)(y)=D_{\mathrm{can}}(ex)(ey)
  =\Omega_{\mathrm{can}}(ex,ey)=\Omega_X(x,y),
\]
so the displayed composition is precisely the induced map.  Each of its three factors is a bounded
linear isomorphism, hence so is $D_X$.
\end{proof}

\begin{theorem}[Canonical normalized-block theorem]\label{thm:block-primary}
For a finite-support sequence $w$, write $\supp w=\{k:w_k\ne0\}$.  Let $(w_n)$ be a
\emph{successive normalized block sequence} in $\ell_2$: every $w_n$ has finite nonempty support,
$\|w_n\|_2=1$, and
\[
  \max\supp w_n<\min\supp w_{n+1}\qquad(n\geq0).
\]
Section~4.4 of Castillo--Cuellar--Gonz\'alez--Pino defines on the $K_0$ model the block operator
$T_w^0$ by
\[
  T_w^0(e_n,0)=(w_n,0),
  \qquad
  T_w^0(0,e_n)=(K_0(w_n),w_n).
\]
Put
\[
  W_w=e_0^{-1}T_w^0e_0
\]
on $Z_2^{\mathrm{can}}$.  Then
\[
  W_w(e_n,0)=(w_n,0),
  \qquad
  W_w(0,e_n)=(\mathrm{KP}(w_n),w_n).
\]
The operator $W_w$ is bounded and injective with complemented range, preserves
$\Omega_{\mathrm{can}}$, and satisfies $W_w^+W_w=I$.  Moreover, $W_wW_w^+$ is a bounded projection
onto $\ran W_w$.  The corresponding Hilbert-space block isometry $U_w:\ell_2\to\ell_2$ satisfies
$U_we_n=w_n$ and intertwines the exact-sequence inclusion:
\[
  W_wi=iU_w.
\]
If $(v_n)$ is another successive normalized block sequence whose supports are disjoint from all
supports of the $w_n$, then
\[
  W_w^+W_v=0=W_v^+W_w.
\]
\uses{thm:kp-canonical-banach,lem:kp-normalization,thm:ks-primary}
\lean{KaltonPeck.Support.Symplectic.IsSuccessiveNormalizedBlockSequence, KaltonPeck.Support.Symplectic.AreMutuallySupportDisjoint, KaltonPeck.Support.Symplectic.canonicalBlockOperator, KaltonPeck.Support.Symplectic.canonicalL2BlockEmbedding, KaltonPeck.Support.Symplectic.canonicalL2BlockEmbedding_norm, KaltonPeck.Support.Symplectic.canonicalL2BlockEmbedding_injective, KaltonPeck.Support.Symplectic.canonicalL2BlockEmbedding_single_apply, KaltonPeck.Support.Symplectic.canonicalBlockOperator_inclusion, KaltonPeck.Support.Symplectic.canonicalNormalizedBlock}
\leanok
\end{theorem}

\begin{proof}
The boundedness, into-isometry, complemented-range, symplectic, and projection assertions for
$T_w^0$ are exactly those stated in \cite[\S4.4]{CCGP}.  Conjugating through $e_0$ gives the
corresponding assertions for $W_w$.  On the two canonical coordinate generators,
\[
 e_0^{-1}T_w^0e_0(e_n,0)
 =e_0^{-1}T_w^0(e_n/2,0)
 =e_0^{-1}(w_n/2,0)
 =(w_n,0)
\]
and
\[
 e_0^{-1}T_w^0e_0(0,e_n)
   =(2K_0(w_n),w_n)
   =(\mathrm{KP}(w_n),w_n).
\]

For every finitely supported $w$, the zero-coordinate convention gives
$\supp K_0(w)\subseteq\supp w$, and likewise $\supp\mathrm{KP}(w)\subseteq\supp w$.  Write
$e_n^0=(e_n,0)$ and $e_n^1=(0,e_n)$.  Mutual support-disjointness gives all four generator
pairings:
\[
\begin{aligned}
 \Omega_{\mathrm{can}}(W_we_n^0,W_ve_m^0)&=0,\\
 \Omega_{\mathrm{can}}(W_we_n^0,W_ve_m^1)&=\langle w_n,v_m\rangle=0,\\
 \Omega_{\mathrm{can}}(W_we_n^1,W_ve_m^0)&=-\langle v_m,w_n\rangle=0,\\
 \Omega_{\mathrm{can}}(W_we_n^1,W_ve_m^1)
   &=\langle\mathrm{KP}(w_n),v_m\rangle
     -\langle\mathrm{KP}(v_m),w_n\rangle=0.
\end{aligned}
\]
Bilinearity gives
\[
  \Omega_{\mathrm{can}}(W_wz,W_vz')=0
\]
for finite coordinate vectors.  Those vectors are dense by
Theorem~\ref{thm:kp-canonical-banach}; continuity of the block maps and of the form extends the
identity to all vectors.  The defining adjoint equation and strongness then give both cross-adjoint
identities.
\end{proof}

\begin{theorem}[Transported normalized-block theorem]\label{thm:block-transport}
Let $X$ be any complete real normed space with a Kalton--Peck presentation, let
$e:X\simeq Z_2^{\mathrm{can}}$ match coordinates, and let $W_w$ be a canonical block operator from
Theorem~\ref{thm:block-primary}.  Put
\[
  \widetilde W_w=e^{-1}W_we.
\]
Then $\widetilde W_w$ is bounded and injective with complemented range, it preserves the
transported form $\Omega_X$, and $\widetilde W_w^+\widetilde W_w=I$.  Moreover,
$\widetilde W_w\widetilde W_w^+$ is a bounded projection onto $\ran\widetilde W_w$.  For a second
successive normalized block sequence $(v_n)$ mutually support-disjoint from $(w_n)$, one also has
\[
  \widetilde W_w^+\widetilde W_v=0
  =\widetilde W_v^+\widetilde W_w.
\]
\uses{thm:block-primary,thm:presentation-equivalence,thm:ks-transport}
\lean{KaltonPeck.Support.Symplectic.transportedBlockOperator, KaltonPeck.Support.Symplectic.transportedNormalizedBlock}
\leanok
\end{theorem}

\begin{proof}
The defining conjugation gives
$e(\widetilde W_wx)=W_w(ex)$.  Transporting the canonical form-preservation identity through
$\Omega_X(x,y)=\Omega_{\mathrm{can}}(ex,ey)$ shows that $\widetilde W_w$ preserves $\Omega_X$.
Strongness and the defining adjoint equation then give
$\widetilde W_w^+\widetilde W_w=I$.  Thus $\widetilde W_w^+$ is a bounded left inverse, which
proves injectivity and complemented range; elementary operator algebra gives the projection and
its range.  Finally, transporting the canonical cross-family identity and applying the adjoint
equation (together with skew-symmetry) yields both cross-adjoint products.
\end{proof}

\begin{definition}[Upper semi-Fredholm]\label{def:upper-semi}
A bounded operator is \emph{upper semi-Fredholm} when its kernel is finite-dimensional and its
range is closed.
\uses{def:fredholm-rank}
\lean{KaltonPeck.Support.GraphFredholm.IsUpperSemiFredholm}
\leanok
\end{definition}

\begin{lemma}[Upper semi-Fredholm composition calculus]\label{lem:upper-semi-calculus}
If $A:X\to Y$ and $B:Y\to Z$ are upper semi-Fredholm operators between real Banach spaces, then
$BA$ is upper semi-Fredholm.
\uses{def:upper-semi}
\lean{KaltonPeck.Support.GraphFredholm.IsUpperSemiFredholm.comp}
\leanok
\end{lemma}

\begin{proof}
The kernel of $BA$ is finite-dimensional because it is an extension of a subspace of
$\ker B$ by $\ker A$.  For closedness of the range, map the closed range of $A$ through the
quotient map $Y\to\ran B$.  Its inverse image is $\ran A+\ker B$, which is closed because
$\ker B$ is finite-dimensional.  The resulting closed subspace of $\ran B$ remains closed in
$Z$.
\end{proof}

\begin{definition}[Infinite-dimensional compact restriction]
\label{def:infinite-dimensional-compact-restriction}
A bounded operator $T:X\to Y$ has an \emph{infinite-dimensional compact restriction} if there
is a closed infinite-dimensional subspace $M\subseteq X$ such that $T|_M$ is compact.
\lean{KaltonPeck.Support.GraphFredholm.HasInfiniteDimensionalCompactRestriction}
\leanok
\end{definition}

\begin{lemma}[The infinite-kernel branch]\label{lem:infinite-kernel-compact-restriction}
If $\ker T$ is infinite-dimensional, then $T$ has an infinite-dimensional compact restriction.
\uses{def:infinite-dimensional-compact-restriction}
\lean{KaltonPeck.Support.GraphFredholm.hasInfiniteDimensionalCompactRestriction_of_infiniteDimensional_kernel}
\leanok
\end{lemma}

\begin{proof}
Take $M=\ker T$.  It is closed, and $T|_M=0$ is compact.
\end{proof}

\begin{lemma}[The nonclosed-range branch]\label{lem:nonclosed-range-approximate-kernel}
Let $T:X\to Y$ have finite-dimensional kernel and nonclosed range.  There are a closed
infinite-dimensional complement $Z$ of $\ker T$ and normalized vectors $z_n\in Z$ for which
\[
  \sum_n \lVert Tz_n\rVert<\infty .
\]
\uses{def:upper-semi}
\lean{KaltonPeck.Support.GraphFredholm.exists_summableApproximateKernelSequence_of_finiteDimensional_ker}
\leanok
\end{lemma}

\begin{proof}
On a closed complement of the kernel, $T$ is injective but not bounded below.  Choose a
nonzero vector on which its norm ratio is less than $2^{-n-1}$ and normalize it.  The geometric
bound makes the image norms summable.  The complement must be infinite-dimensional, since a
finite-dimensional domain would have closed image.
\end{proof}

\begin{lemma}[Failure of upper semi-Fredholmness]\label{lem:not-upper-semi-dichotomy}
If $T$ is not upper semi-Fredholm, then either $T$ has an infinite-dimensional compact
restriction or the summable normalized approximate-kernel sequence of
Lemma~\ref{lem:nonclosed-range-approximate-kernel} exists on a closed kernel complement.
\uses{def:upper-semi,def:infinite-dimensional-compact-restriction,lem:infinite-kernel-compact-restriction,lem:nonclosed-range-approximate-kernel}
\lean{KaltonPeck.Support.GraphFredholm.compactRestriction_or_summableApproximateKernelSequence}
\leanok
\end{lemma}

\begin{proof}
Split according to whether the kernel is finite-dimensional and apply the preceding two lemmas.
Upgrading the second alternative to a compact restriction requires a basic-sequence selection
theorem; that gliding-hump step is not part of this lemma.
\end{proof}

\begin{lemma}[Controlled approximation by the canonical kernel]
\label{lem:canonical-kernel-approximation}
If $q:X\to Y$ is a surjective bounded operator between Banach spaces, then some $C>0$ satisfies
\[
  \inf_{y\in\ker q}\lVert x-y\rVert\le C\lVert qx\rVert
\]
for every $x\in X$.  In particular this holds for the canonical quotient
$p:Z_2^{\mathrm{can}}\to\ell_2$.
\uses{thm:kp-canonical-banach}
\lean{KaltonPeck.Support.GraphFredholm.exists_kernel_approximation_of_surjective, KaltonPeck.Support.GraphFredholm.canonicalL2Quotient_kernel_approximation}
\leanok
\end{lemma}

\begin{proof}
The open mapping theorem supplies a preimage $z$ of $qx$ with
$\lVert z\rVert\le C\lVert qx\rVert$.  Then $x-z\in\ker q$, so $z$ is the required error.
\end{proof}

\begin{lemma}[Compact canonical block extraction]\label{lem:compact-canonical-block-extraction}
Let $X$ be a real Banach space and $T:\ell_2\to X$ a bounded operator which is not upper
semi-Fredholm.  There is a successive normalized finite-support block family $(w_n)$ such that
\[
  \sum_n\lVert TU_we_n\rVert<\infty.
\]
Consequently, the operator $TU_w$ is compact.
\uses{def:upper-semi,thm:block-primary}
\lean{KaltonPeck.Support.CgpBlockExtraction.exists_summable_canonicalL2Block_of_not_upperSemi, KaltonPeck.Support.CgpBlockExtraction.exists_compact_canonicalL2Block_of_not_upperSemi}
\leanok
\end{lemma}

\begin{proof}
For every finite coordinate head, $T$ fails to be bounded below on the orthogonal tail;
otherwise the kernel embeds into the head and the range is the sum of a closed tail image and a
finite-dimensional space.  Choose successively normalized tail vectors with geometrically small
images, approximate them by finite truncations, and arrange disjoint support intervals.  The
images of the standard basis under $TU_w$ then have summable norms.  Expanding $TU_w$ as a
norm-convergent sum of rank-one maps proves compactness.
\end{proof}

\begin{lemma}[Hilbert gliding-hump approximation]\label{lem:hilbert-gliding-hump}
Every infinite-dimensional subspace $M\subseteq\ell_2$ contains unit vectors $(y_n)$ which are
summably close to a successive normalized finite-support block sequence $(w_n)$.  Every continuous
functional on $\ell_2$ tends to zero along the standard unit-vector basis.  Moreover, every
normalized coordinatewise-null sequence $(u_n)$ has a strictly increasing subsequence
$(u_{\phi(n)})$ summably close to a successive normalized finite-support block sequence.
In particular, all normalized signed averages of the first family $(y_n)$ are bounded uniformly by
one plus the total approximation error.
\uses{thm:block-primary}
\lean{KaltonPeck.Support.HilbertGlidingHump.exists_summable_successiveBlock_approximation, KaltonPeck.Support.HilbertGlidingHump.tendsto_dual_canonicalL2Basis_zero, KaltonPeck.Support.HilbertGlidingHump.exists_summable_successiveBlock_subsequence_approximation, KaltonPeck.Support.HilbertGlidingHump.norm_signed_average_le_one_add_tsum_error, KaltonPeck.Support.HilbertGlidingHump.exists_successiveBlock_with_uniform_signed_average_bound}
\leanok
\end{lemma}

\begin{proof}
Kill an arbitrary finite coordinate head inside $M$, truncate the resulting unit vector far
enough out, normalize, and iterate on successive support intervals with geometric errors.  The
exact block vectors are orthonormal, so the triangle inequality gives the signed-average bound.
For the subsequence form, recursively pass far enough along the coordinatewise-null sequence that
its finite coordinate head is small, truncate the tail, and normalize with geometrically summable
errors.  The selected indices are strictly increasing and the truncation intervals are successive.
The assertion about continuous functionals follows from the Riesz representation theorem and the
fact that every square-summable coordinate sequence tends to zero.
\end{proof}

\begin{definition}[Bounded centralizer data]\label{def:bounded-centralizer-data}
A bounded centralizer lift of $Q:Z\to\ell_2$ is a linear sequence-valued map whose difference
from the Kalton--Peck centralizer of $Qz$ is square-summable with norm bounded by
$C\lVert z\rVert$.  A bounded centralizer approximation on a subspace
$M\subseteq\ell_2$ is the corresponding datum for the inclusion of $M$.
\uses{thm:kp-canonical-banach}
\lean{KaltonPeck.Support.GraphFredholm.HasBoundedCentralizerLift, KaltonPeck.Support.GraphFredholm.HasBoundedCentralizerApproximation}
\leanok
\end{definition}

\begin{lemma}[Canonical quotient reduction]
\label{lem:canonical-quotient-strictly-singular-reduction}
Every bounded map $S:Z\to Z_2^{\mathrm{can}}$ gives a bounded centralizer lift of $pS$, where
$p:Z_2^{\mathrm{can}}\to\ell_2$ is the canonical quotient.  Consequently, if no closed
infinite-dimensional subspace of $\ell_2$ admits a bounded centralizer approximation, then $p$
is strictly singular.
\uses{thm:kp-canonical-banach,def:strictly-singular,def:bounded-centralizer-data}
\lean{KaltonPeck.Support.GraphFredholm.hasBoundedCentralizerLift_canonicalL2Quotient_comp, KaltonPeck.Support.GraphFredholm.no_boundedCentralizerLift_of_subspace_obstruction, KaltonPeck.Support.GraphFredholm.canonicalL2Quotient_strictlySingular_of_centralizer_obstruction, KaltonPeck.Support.GraphFredholm.canonicalL2Quotient_strictlySingular_of_subspace_obstruction}
\leanok
\end{lemma}

\begin{proof}
Take the first coordinate of $Sz$ as the linear lift.  Membership in the canonical model gives
square-summability of its centralizer defect, and the norm comparison bounds that defect by
$\lVert S\rVert\lVert z\rVert$.  If $Q$ is bounded below, its range is a closed
infinite-dimensional Hilbert subspace and the inverse on that range transports the lift to a
bounded centralizer approximation.  The asserted obstruction therefore rules out every
bounded-below restriction of $p$.
\end{proof}

\begin{lemma}[Summable weighted Schauder images are compact]
\label{lem:schauder-weighted-images-compact}
Let $(x_n)$ be a Schauder basis of a Banach space $X$, with coordinate functionals $x_n^*$.
If $T:X\to Y$ is bounded and
\[
  \sum_n \lVert x_n^*\rVert\lVert Tx_n\rVert<\infty,
\]
then $T$ is compact.
\lean{KaltonPeck.Support.CgpBlockExtraction.compact_of_schauderBasis_weighted}
\leanok
\end{lemma}

\begin{proof}
The rank-one series $\sum_n x_n^*\otimes Tx_n$ converges in operator norm by the displayed
bound.  Its partial sums are compact and its pointwise sum equals $T$ by the Schauder expansion,
so closure of the compact operators in operator norm gives the result.
\end{proof}

\begin{lemma}[Basic-sequence compact-restriction bridge]
\label{lem:basic-sequence-compact-restriction}
If a bounded operator between Banach spaces is not upper semi-Fredholm, then it is compact on
some closed infinite-dimensional subspace of its domain.
\uses{def:upper-semi,def:infinite-dimensional-compact-restriction,lem:not-upper-semi-dichotomy,lem:schauder-weighted-images-compact}
\lean{KaltonPeck.Support.GraphFredholm.exists_biorthogonal_sequence_summable_of_not_upper, KaltonPeck.Support.GraphFredholm.hasInfiniteDimensionalCompactRestriction_of_biorthogonal_weighted, KaltonPeck.Support.GraphFredholm.hasInfiniteDimensionalCompactRestriction_of_not_upper}
\leanok
\end{lemma}

\begin{proof}
Recursively choose normalized vectors $x_n$ and continuous functionals $x_n^*$ with
$x_i^*(x_j)=\delta_{ij}$, making
\[
  \sum_n \lVert x_n^*\rVert\lVert Tx_n\rVert
\]
finite.  At each stage the common kernel of the previously chosen functionals has finite
codimension; a lower bound for $T$ there would force finite-dimensional kernel and closed range,
so failure of upper semi-Fredholmness supplies the next small image.  The closed span of the
$x_n$ is infinite-dimensional by biorthogonality.  On that span, the resulting rank-one
expansion of $T$ is the norm limit of finite-rank operators, and hence is compact.
\end{proof}

\begin{theorem}[Canonical centralizer obstruction]
\label{thm:canonical-centralizer-obstruction}
No closed infinite-dimensional subspace of $\ell_2$ admits a bounded linear approximation to the
Kalton--Peck centralizer.  Consequently the canonical quotient
$p:Z_2^{\mathrm{can}}\to\ell_2$ is strictly singular.
\uses{def:bounded-centralizer-data,lem:canonical-quotient-strictly-singular-reduction,lem:canonical-kernel-approximation,lem:hilbert-gliding-hump,thm:block-primary}
\lean{KaltonPeck.Support.GraphFredholm.no_boundedCentralizerApproximation, KaltonPeck.Support.GraphFredholm.canonicalL2Quotient_isStrictlySingular}
\leanok
\end{theorem}

\begin{proof}
Assume a bounded approximation on $M$ and bundle it as a bounded lift $L:M\to Z_2^{\mathrm{can}}$.
Apply Lemma~\ref{lem:hilbert-gliding-hump}, compare $Ly_n$ with the corresponding canonical block
images, and approximate their differences by vectors in $\ker p=i[\ell_2]$.  A Hilbert-space
sign choice keeps the normalized signed kernel errors uniformly bounded.  Hence all signed
canonical block images remain uniformly bounded.  Their preimages under the block operator have
exact quasi-norm $\log N+1$, contradicting the bounded block left inverse as $N\to\infty$.
\end{proof}

\begin{lemma}[Corrected canonical upper-restriction implication]
\label{lem:corrected-cgp-upper-restriction}
Let $B$ be a bounded operator on $Z_2^{\mathrm{can}}$.  If $B$ is not upper
semi-Fredholm, then neither is its kernel restriction $Bi:\ell_2\to Z_2^{\mathrm{can}}$.
Moreover, there is a successive normalized Hilbert block embedding $U_w$ for which
\[
  BiU_w
\]
is compact.  Equivalently, the target-minimal canonical kernel compact-block extraction
property holds.
\uses{lem:basic-sequence-compact-restriction,lem:canonical-kernel-approximation,lem:compact-canonical-block-extraction,thm:canonical-centralizer-obstruction,lem:upper-semi-not-strictly-singular}
\lean{KaltonPeck.Support.GraphFredholm.Canonical.exists_compact_kernel_embedding_of_compactRestriction, KaltonPeck.Support.GraphFredholm.Canonical.not_upperSemi_comp_canonicalL2Inclusion_of_not_upper, KaltonPeck.Support.GraphFredholm.Canonical.exists_compact_canonicalL2Block_of_not_upper, KaltonPeck.Support.GraphFredholm.canonical_not_upperSemi_comp_canonicalL2Inclusion_of_not_upper, KaltonPeck.Support.GraphFredholm.HasCanonicalKernelCompactBlockExtraction, KaltonPeck.Support.GraphFredholm.hasCanonicalKernelCompactBlockExtraction}
\leanok
\end{lemma}

\begin{proof}
By Lemma~\ref{lem:basic-sequence-compact-restriction}, $B$ is compact on a closed
infinite-dimensional subspace $M$.  Strict singularity of $p$ yields a biorthogonal sequence in
$M$ on which $p$ is summably small.  Approximate those vectors by vectors in $\ker p$ and sum the
resulting rank-one errors to obtain a compact operator $K$ with $\lVert K\rVert<1$ on a tail.
The map $I-K$ is bounded below there, its image lies in $\ker p=\ran i$, and $B(I-K)$ is compact.
Factoring this embedding through $i$ gives a bounded-below map $R$ for which $BiR$ is compact.
If $Bi$ were upper semi-Fredholm, then so would be $BiR$, contradicting strict singularity of a
compact operator on an infinite-dimensional space.  Finally apply
Lemma~\ref{lem:compact-canonical-block-extraction} to $Bi$.
\end{proof}

\begin{lemma}[Canonical exact-sequence adjoint bridges]
\label{lem:canonical-exact-sequence-adjoint-bridges}
Write $i:\ell_2\to Z_2^{\mathrm{can}}$ and
$p:Z_2^{\mathrm{can}}\to\ell_2$ for the canonical inclusion and quotient.  The canonical form
satisfies
\[
  \Omega_{\mathrm{can}}(ix,z)=\langle x,pz\rangle,
  \qquad
  \Omega_{\mathrm{can}}(z,ix)=-\langle pz,x\rangle .
\]
If $g_k$ is the $k$th canonical quotient-coordinate basis vector, then
\[
  \Omega_{\mathrm{can}}(ix,g_k)=x_k.
\]
Furthermore, if $(w_n)$ is a successive normalized Hilbert block sequence, $U_w$ its block
isometry, $A:\ell_2\to Z_2^{\mathrm{can}}$ is bounded, and $z$ is fixed, then
\[
  \Omega_{\mathrm{can}}(AU_we_n,z)\longrightarrow0.
\]
If $Ri=0$, then $R=Lp$ for a bounded $L:\ell_2\to Z_2^{\mathrm{can}}$, and $R^+$ has range in
$\ker p=\ran i$, hence factors through $i$.  More precisely, if $Vi=iv$, then
\[
  (R^+V)i=i\,(pL)^*v.
\]
For every bounded $T$ on the canonical space,
\[
  pT^+i=-(pTi)^*.
\]
Both $pTi$ and $pT^+i$ are therefore strictly singular.
\uses{thm:ks-primary,thm:block-primary,lem:hilbert-gliding-hump,lem:strictly-singular-comp,thm:canonical-centralizer-obstruction}
\lean{KaltonPeck.Support.Symplectic.canonicalKaltonSwansonForm_inclusion_left, KaltonPeck.Support.Symplectic.canonicalKaltonSwansonForm_inclusion_right, KaltonPeck.Support.GraphFredholm.canonical_pairing_inclusion_secondBasis, KaltonPeck.Support.GraphFredholm.tendsto_canonical_pairing_apply_block_zero, KaltonPeck.Support.GraphFredholm.canonical_factor_through_quotient, KaltonPeck.Support.GraphFredholm.canonical_symplecticAdjoint_range_le_quotientKernel, KaltonPeck.Support.GraphFredholm.canonical_symplecticAdjoint_factor_through_inclusion, KaltonPeck.Support.GraphFredholm.canonical_remainderAdjoint_kernel_formula, KaltonPeck.Support.GraphFredholm.canonical_symplecticAdjoint_kernelCompression, KaltonPeck.Support.GraphFredholm.canonical_quotient_operator_inclusion_isStrictlySingular, KaltonPeck.Support.GraphFredholm.canonical_quotient_symplecticAdjoint_inclusion_isStrictlySingular}
\leanok
\end{lemma}

\begin{proof}
The two pairing formulas follow first on finite-coordinate vectors and then by density and
continuity.  The basis-coordinate identity is the first pairing formula specialized to $g_k$.
For the limit identity, composition with $A$, $U_w$, and pairing against $z$ gives a continuous
functional on $\ell_2$; apply the weak convergence of the standard unit-vector basis.  The
quotient universal property factors every operator which kills $\ran i$ through $p$.  Applying
the defining adjoint equation to the first pairing formula puts the adjoint range
in $\ker p$, where the inverse of $i$ gives the second factorization.  The same pairing calculation
identifies the displayed cross term and compression with the ordinary Hilbert adjoints.  Finally,
strict singularity of $p$, composition closure, Hilbert-adjoint invariance, and negation closure
give the last assertion.
\end{proof}

\begin{lemma}[Canonical strictly-singular lifting]
\label{lem:canonical-strictly-singular-lifting}
Let $T$ be a bounded operator on $Z_2^{\mathrm{can}}$.  If its restriction
$Ti:\ell_2\to Z_2^{\mathrm{can}}$ is strictly singular, then $T$ is strictly singular.
More generally, this conclusion follows from strict singularity of the canonical quotient $p$
and of $Ti$.
\uses{def:strictly-singular,lem:strictly-singular-comp,lem:canonical-kernel-approximation,thm:canonical-centralizer-obstruction}
\lean{KaltonPeck.Support.GraphFredholm.canonical_isStrictlySingular_of_inclusion_of_quotient, KaltonPeck.Support.GraphFredholm.canonical_isStrictlySingular_of_inclusion}
\leanok
\end{lemma}

\begin{proof}
Suppose instead that an infinite-dimensional Banach space embeds through a bounded-below map
$S$ so that $TS$ is also bounded below.  Since $pS$ is strictly singular, choose a biorthogonal
sequence on which its weighted image norms are summable.  On a sufficiently far tail,
Lemma~\ref{lem:canonical-kernel-approximation} gives kernel vectors close enough to the images
under $S$.  A norm-convergent rank-one interpolation series produces a small operator $P$
matching these errors.

On the closed span of the tail, $J=S-P$ takes values in $\ker p=\ran i$.  The perturbation is
small enough that both $J$ and $TJ$ remain bounded below.  Factor $J=iL$ through the canonical
inclusion.  Then $L$ is bounded below and $(Ti)L=TJ$ is bounded below, contradicting strict
singularity of $Ti$.  Strict singularity of $p$ is supplied by
Theorem~\ref{thm:canonical-centralizer-obstruction}.
\end{proof}

\begin{theorem}[Canonical CGP operator propositions]
\label{thm:cgp-operator-propositions}
For every bounded $T:Z_2^{\mathrm{can}}\to Z_2^{\mathrm{can}}$:
\begin{enumerate}
\item if $T^+T$ is strictly singular, then $T$ is strictly singular;
\item if $Ti$ is strictly singular for the canonical inclusion $i:\ell_2\to Z_2^{\mathrm{can}}$,
then $T$ is strictly singular;
\item if $Ti$ is upper semi-Fredholm, then $T$ is upper semi-Fredholm.
\end{enumerate}
\uses{def:upper-semi,def:strictly-singular,lem:strictly-singular-comp,lem:canonical-hilbert-strictly-singular-compact,lem:canonical-kernel-approximation,lem:basic-sequence-compact-restriction,lem:corrected-cgp-upper-restriction,lem:canonical-exact-sequence-adjoint-bridges,lem:canonical-strictly-singular-lifting,thm:canonical-centralizer-obstruction,thm:block-primary}
\notready
\end{theorem}

\begin{proof}
These are CGP Proposition~5.3(a)--(c).  Part (b) is
Lemma~\ref{lem:canonical-strictly-singular-lifting}.  Part (c), including the corrected
published biorthogonal perturbation argument on $(I-K)[N]$, is
Lemma~\ref{lem:corrected-cgp-upper-restriction}.  The additive and adjoint-stability layers, the
Hilbert-space equivalence between strict singularity and compactness, and the exact-sequence
bridges required by the source strategy are also formalized.  The full three-part node remains
open only because part (a), the implication from strict singularity of $T^+T$ to strict
singularity of $T$, has not yet been derived.
\end{proof}

\begin{lemma}[Compact biadjoint kernel-column correction]
\label{lem:compact-biadjoint-kernel-correction}
Let $(d_n)$ be a sequence in $Z_2^{\mathrm{can}}$ satisfying
\[
  \sum_n\lVert d_n\rVert<\infty.
\]
There is a compact operator $K$ on $Z_2^{\mathrm{can}}$ whose symplectic adjoint $K^+$ is also
compact and which sends the canonical kernel basis vector $f_n=i(e_n)$ to $d_n$ for every $n$.
\uses{thm:ks-primary,lem:canonical-exact-sequence-adjoint-bridges}
\lean{KaltonPeck.Support.GraphFredholm.exists_compact_biadjoint_kernelCorrection}
\leanok
\end{lemma}

\begin{proof}
Pairing against the canonical second-coordinate basis gives uniformly bounded functionals
$\varphi_n$ satisfying $\varphi_n(f_m)=\delta_{nm}$.  The rank-one series
\[
  K=\sum_n \varphi_n\otimes d_n
\]
converges in operator norm, interpolates the prescribed values, and is compact.  Computing
symplectic adjoints term by term gives another norm-convergent series of rank-one operators:
its $n$th term sends $z$ to $\Omega_{\mathrm{can}}(d_n,z)$ times the $n$th
second-coordinate basis vector.  Hence $K^+$ is compact as well.
\end{proof}

\begin{definition}[Upper summable kernel-block approximation]
\label{def:upper-summable-kernel-block-approximation}
The \emph{upper summable kernel-block approximation property} asserts that every upper
semi-Fredholm operator $T$ admits canonical normalized block operators $W_w,W_v$ and a nonzero
real scalar $\alpha$ for which
\[
  \sum_n\left\lVert T W_w f_n-\alpha W_v f_n\right\rVert<\infty,
  \qquad f_n=i(e_n).
\]
\uses{def:upper-semi,thm:block-primary}
\lean{KaltonPeck.Support.GraphFredholm.HasUpperSummableKernelBlockApproximation}
\leanok
\end{definition}

\begin{lemma}[Canonical upper summable kernel-block selection]
\label{lem:upper-summable-kernel-block-selection}
The upper summable kernel-block approximation property holds on
$Z_2^{\mathrm{can}}$.
\uses{def:upper-summable-kernel-block-approximation,lem:upper-semi-calculus,lem:canonical-hilbert-strictly-singular-compact,lem:upper-semi-not-strictly-singular,lem:compact-canonical-block-extraction,lem:canonical-kernel-approximation,lem:hilbert-gliding-hump,lem:canonical-exact-sequence-adjoint-bridges,thm:canonical-centralizer-obstruction,thm:block-primary}
\lean{KaltonPeck.Support.GraphFredholm.hasUpperSummableKernelBlockApproximation}
\leanok
\end{lemma}

\begin{proof}
Let $T$ be upper semi-Fredholm and put $A=Ti$.  The quotient compression $pA$ is strictly
singular and hence compact on $\ell_2$; the orthogonal projection onto $\ker A$ is compact as well.
Apply the summable block-extraction theorem to their product.  This gives normalized successive
source blocks $(b_n)$ for which both quotient images and kernel projections are summable.

Closed range and finite-dimensional kernel give a lower bound for $A$ on the orthogonal complement
of its kernel.  After discarding a finite head, the small kernel projections therefore make
$\lVert Ab_n\rVert$ uniformly positive.  Controlled approximation by $\ker p=\ran i$ supplies
vectors $a_n\in\ell_2$ such that $Ab_n-ia_n$ is summable, while the norms $\lVert a_n\rVert$ stay
between two positive constants.  Pass to a subsequence whose norms converge summably fast to some
$\alpha>0$.

The canonical pairing identities show that every coordinate of $a_n$ tends to zero.  Normalize and
apply the coordinatewise-null gliding-hump theorem to obtain successive target blocks.  Combining
the kernel-approximation errors, the scalar norm errors, and the target block-approximation errors
gives
\[
  \sum_n\lVert TW_wf_n-\alpha W_vf_n\rVert<\infty ,
\]
which is the required property.
\end{proof}

\begin{definition}[Retained upper compact-Gram factorization]
\label{def:upper-compact-gram-factorization}
The \emph{upper compact-Gram factorization property} asserts that every upper semi-Fredholm
$T$ admits bounded operators $W,V,R,K$ on $Z_2^{\mathrm{can}}$, bounded operators $u,v$ on
$\ell_2$, and a nonzero real scalar $\alpha$ such that
\[
\begin{gathered}
  Wi=iu,\qquad Vi=iv,\qquad V^+V=I,\\
  TW=\alpha V+R+K,\qquad Ri=0,
\end{gathered}
\]
and both $K$ and $K^+$ are compact.
\uses{def:upper-semi,thm:ks-primary}
\lean{KaltonPeck.Support.GraphFredholm.HasUpperCompactGramFactorization}
\leanok
\end{definition}

\begin{lemma}[Summable kernel blocks give the retained factorization]
\label{lem:summable-kernel-block-factorization}
The upper summable kernel-block approximation property implies the retained upper compact-Gram
factorization property.
\uses{def:upper-summable-kernel-block-approximation,def:upper-compact-gram-factorization,lem:compact-biadjoint-kernel-correction,thm:block-primary}
\lean{KaltonPeck.Support.GraphFredholm.hasUpperCompactGramFactorization_of_summableKernelBlockApproximation}
\leanok
\end{lemma}

\begin{proof}
For an upper semi-Fredholm $T$, choose $W=W_w$, $V=W_v$, and $\alpha\ne0$ from the summable
block approximation.  Apply Lemma~\ref{lem:compact-biadjoint-kernel-correction} to the errors
$d_n=TWf_n-\alpha Vf_n$, obtaining $K$ with both $K$ and $K^+$ compact.  Set
\[
  R=TW-\alpha V-K.
\]
The interpolation identities give $Rf_n=0$ for every canonical kernel basis vector.  Density of
their span and continuity imply $Ri=0$.  The normalized-block identities give
$Wi=iu$, $Vi=iv$, and $V^+V=I$, so these operators satisfy every clause of the retained
factorization.
\end{proof}

\begin{lemma}[Target-specific compact-Gram reduction]
\label{lem:target-compact-gram-reduction}
Assume the upper compact-Gram factorization property.  If $T$ is upper semi-Fredholm, then the
kernel restriction $(T^+T)i$ is not compact.  Together with the corrected canonical
compact-block extraction, this gives
\[
  A\ \hbox{upper semi-Fredholm}\quad\Longrightarrow\quad A^+A\
  \hbox{upper semi-Fredholm}.
\]
\uses{def:upper-compact-gram-factorization,lem:canonical-exact-sequence-adjoint-bridges,lem:corrected-cgp-upper-restriction,lem:strictly-singular-comp,lem:compact-strictly-singular,lem:upper-semi-not-strictly-singular,thm:block-primary}
\lean{KaltonPeck.Support.GraphFredholm.HasCanonicalKernelCompactBlockExtraction, KaltonPeck.Support.GraphFredholm.upperSemi_gramKernel_not_compact_of_factorization, KaltonPeck.Support.GraphFredholm.canonicalGram_isUpperSemiFredholm_of_compactBlock_and_noncompactGram, KaltonPeck.Support.GraphFredholm.canonicalGram_isUpperSemiFredholm_of_targetFactorizations}
\leanok
\end{lemma}

\begin{proof}
Conjugate the assumed compact restriction of $T^+T$ by $W$.  Expanding the Gram operator of
$TW=\alpha V+R+K$ leaves the leading term $\alpha^2 i$.  The terms containing $K$ are compact
and hence strictly singular.  Since $R=Lp$ and $Vi=iv$, the only remaining cross term is
$i(pL)^*v$, which is strictly singular by
Lemma~\ref{lem:canonical-exact-sequence-adjoint-bridges}.  Additive ideal calculus would therefore
make the nonzero scalar multiple $\alpha^2 i$ strictly singular, contradicting upper
semi-Fredholmness of $i$.

For the consequence, suppose $B=A^+A$ is not upper semi-Fredholm and extract a block operator
$W$ for which $BWi$ is compact.  Then $AW$ is upper semi-Fredholm, while its Gram restriction
$W^+BWi$ is compact, contradicting the preceding noncompactness conclusion.
\end{proof}

\begin{lemma}[Formal CGP Gram reduction]\label{lem:cgp-primary-reduction}
Assume the three implications in Theorem~\ref{thm:cgp-operator-propositions}.  If $A$ is upper
semi-Fredholm on $Z_2^{\mathrm{can}}$, then $A^+A$ is upper semi-Fredholm.  Every canonical
normalized block operator is itself upper semi-Fredholm.
\uses{thm:cgp-operator-propositions,def:upper-semi,lem:upper-semi-calculus,lem:strictly-singular-comp,lem:compact-strictly-singular,lem:upper-semi-not-strictly-singular,lem:compact-canonical-block-extraction,thm:block-primary}
\lean{KaltonPeck.Support.GraphFredholm.canonicalBlockOperator_isUpperSemiFredholm, KaltonPeck.Support.GraphFredholm.canonicalGram_isUpperSemiFredholm_of_cgpPropositions}
\leanok
\end{lemma}

\begin{proof}
Put $B=A^+A$ and suppose $B$ is not upper semi-Fredholm.  Proposition~5.3(c) implies that $Bi$
is not upper semi-Fredholm, so choose a Hilbert block embedding $i_w$ for which $Bii_w$ is
compact.  For its canonical block operator $W$, the identity $Wi=ii_w$ makes $(BW)i$ compact,
hence strictly singular.  Proposition~5.3(b) makes $BW$ strictly singular, and the
strictly-singular composition calculus gives
\[
  (AW)^+(AW)=W^+BW
\]
strictly singular.  Proposition~5.3(a) makes $AW$ strictly singular.  But $W^+W=I$, so $W$ is
upper semi-Fredholm; composition calculus makes $AW$ upper semi-Fredholm, a contradiction.
\end{proof}

\begin{theorem}[Canonical Castillo--Gonz\'alez--Pino theorem]\label{thm:cgp-primary}
On $Z_2^{\mathrm{can}}$ with the canonical Kalton--Swanson form, let $A$ be a bounded operator.  If
$\ker A$ is finite-dimensional and $\ran A$ is closed, then $A^+A$ is Fredholm.
\uses{thm:ks-primary,def:fredholm-rank,lem:upper-summable-kernel-block-selection,lem:summable-kernel-block-factorization,lem:target-compact-gram-reduction,lem:adjoint-calculus}
\lean{KaltonPeck.Support.GraphFredholm.cgpPrimary}
\leanok
\end{theorem}

\begin{proof}
Lemma~\ref{lem:upper-summable-kernel-block-selection} establishes the required block-selection
property.  Lemma~\ref{lem:summable-kernel-block-factorization} therefore supplies the retained
factorization, and
Lemma~\ref{lem:target-compact-gram-reduction} makes $A^+A$ upper semi-Fredholm.
The operator $A^+A$ is self-adjoint.  On a strongly symplectic Banach space the kernel of a
self-adjoint operator is isomorphic to the annihilator of its range, so finite-dimensional
kernel and closed range force finite-dimensional cokernel.  Hence $A^+A$ is Fredholm.
\end{proof}

\begin{theorem}[Transported Castillo--Gonz\'alez--Pino theorem]\label{thm:cgp-transport}
Let $X$ be a complete real normed space with a Kalton--Peck presentation and its transported
Kalton--Swanson form.  If a bounded operator $A$ on $X$ is upper semi-Fredholm, then $A^+A$ is
Fredholm.
\uses{thm:cgp-primary,thm:presentation-equivalence,thm:ks-transport,lem:fredholm-calculus,lem:adjoint-calculus}
\lean{KaltonPeck.Support.GraphFredholm.cgpTransport}
\leanok
\end{theorem}

\begin{proof}
Let $e:X\simeq Z_2^{\mathrm{can}}$ match coordinates and set $A_{\mathrm{can}}=eAe^{-1}$.  The map
$e$ identifies $\ker A$ with $\ker A_{\mathrm{can}}$ and $\ran A$ with
$\ran A_{\mathrm{can}}$; because $e$ is a homeomorphism, it also preserves finite-dimensionality and
closedness.  Hence $A_{\mathrm{can}}$ is upper semi-Fredholm, so
$A_{\mathrm{can}}^+A_{\mathrm{can}}$ is Fredholm by Theorem~\ref{thm:cgp-primary}.

The formula defining the transported form gives
\[
  A_{\mathrm{can}}^+=eA^+e^{-1},
  \qquad
  A_{\mathrm{can}}^+A_{\mathrm{can}}=e(A^+A)e^{-1}.
\]
The first identity follows by inserting $eAe^{-1}$ into the defining adjoint equation and using
uniqueness from strongness.  Fredholm invariance under bounded linear equivalence transports the
conclusion back to $A^+A$.
\end{proof}
